Hire Ed: Job Market Dynamics for Tenure-Track Faculty Positions in Archaeology

### 2025-05-14 analysis/paper/article.qmd
fix a citation

Longer-term editors who represent job seekers active for multiple seasons reflect one of the underdiscussed realities of the academic job market in archaeology—time spent on that market. Previous research on professional trajectories in academic archaeology suggests that early career researchers have a limited shelf life. Mackie and Rockwell, for example, identify a hiring plateau that occurs around seven years after candidates defend their PhDs [-@mackieMarketShareAccounting2023:5]. However, too little time on the market also seems to have a negative effect on placement rates—“very few graduates obtain TT employment when they are all but dissertation (ABD) or even immediately after graduation” [@mackieMarketShareAccounting2023:5]. These patterns are not restricted to the United States. In Brami et al.’s [-@bramiPrecariousFutureReflections2023] survey of 419 early career researchers in European archaeology, likelihood of holding a permanent position shifted relative to time since PhD. Of the respondents who were 1 year or less to 4 years post-PhD, only 3% (3/116) held permanent positions. These numbers increased slightly for participants who had been on the market for 5–7 years, with 14% (7/49) holding permanent positions; similarly, 18% (6/34) of respondents who were on the market for 8 years or more held permanent positions.

>>>>

Longer-term editors who represent job seekers active for multiple seasons reflect one of the underdiscussed realities of the academic job market in archaeology—time spent on that market. Previous research on professional trajectories in academic archaeology suggests that early career researchers have a limited shelf life. Mackie and Rockwell, for example, identify a hiring plateau that occurs around seven years after candidates defend their PhDs [-@mackieMarketShareAccounting2023, p. 5]. However, too little time on the market also seems to have a negative effect on placement rates—"very few graduates obtain TT employment when they are all but dissertation (ABD) or even immediately after graduation" [@mackieMarketShareAccounting2023, p. 5]. These patterns are not restricted to the United States. In Brami et al.’s [-@bramiPrecariousFutureReflections2023] survey of 419 early career researchers in European archaeology, likelihood of holding a permanent position shifted relative to time since PhD. Of the respondents who were 1 year or less to 4 years post-PhD, only 3% (3/116) held permanent positions. These numbers increased slightly for participants who had been on the market for 5–7 years, with 14% (7/49) holding permanent positions; similarly, 18% (6/34) of respondents who were on the market for 8 years or more held permanent positions.
